Interfacing logic programming and Coq

### 2020-10-24 papers/types2020/sec3.tex
adding refs for LJT

\begin{metanote}
  (AM): I would start with the logical reconstruction of \texttt{check}, adding the LJT-style proof terms to fig \ref{fig:augmented} and then reverse engineer what we want/need to say about focusing and FPC. We \textbf{do} not talk about PBT yet, just logic programming (with proof terms). In sec 4 we package \texttt{check} in a PBT query and explain how \texttt{check} works for term generation.
\end{metanote}
\subsection{Focused proof systems}
\label{ssec:focused}

The proof search approach to encoding Horn clause computation
results in the structuring of proofs with repeated
switching between a \emph{goal-reduction} phase and a
\emph{backchaining} phase~\cite{miller91apal}.
%
The notion of \emph{focused proof systems}~\cite{andreoli92jlc,liang09tcs}
generalizes this view of proof construction by identifying the
following two phases. % of proof construction.
\begin{enumerate}
\item The \emph{negative}%
\footnote{The terminology of negative and positive phases is a bit
  unfortunate: historically, these terms do not refer to positive or negative
  subformula occurrences but rather to certain semantic models used in
  the study of linear logic~\cite{girard87tcs}.}
%
  phase corresponds to  goal-reduction: in this phase,
  inference rules that involve \emph{don't-care-non\-de\-ter\-min\-ism} are
  applied.
%
  As a result, there is no need to consider backtracking over choices
  made in building this phase.

@@

Further, if we view a particular FPC as a means of restricting proofs, it is
possible to build an FPC that \emph{restricts} proofs satisfying two FPCs
simultaneously.
%
In particular, Figure~\ref{fig:pairing} defines an FPC based on the
(infix) constructor \lsti{<c>}, which \emph{pairs} two terms of type
\lsti{cert}.
%
The pairing experts for the certificate \lsti{Cert1 <c> Cert2} simply
request that the corresponding experts also succeed for both
\lsti{Cert1} and \lsti{Cert2} and, in the case of the \lsti{orE} and
\lsti{someE}, also return the same choice and substitution term,
respectively.
%
Thus, the query 
\begin{lstlisting}
?- check ((qheight 4) <c> (qsize 10)) B
\end{lstlisting}
will succeed if there is a proof of \lsti{B} that has a height less
than or equal to 4 while also being of size less than or equal to 10.
%
% A related use of the pairing of two proof certificates is to %use it to
% \emph{distill} or \emph{elaborate} proof certificates. 
% %
% For example, the proof certificate \lsti{(qsize 5 0)} is rather implicit
% since it will match any proof that used unfold exactly 5 times.
% %
% However, the query
% \begin{lstlisting}
% ?- check ((qsize 5 0) <c> (max Max)) B.
% \end{lstlisting}
% will store into the \lP variable \lsti{Max} more complete details of
% any proof that satisfies the \lsti{(qsize 5 0)} constraint.
% %
% In particular, this forms the infrastructure of 
%  an \emph{explanation} tool for attributing ``blame'' for
% the origin of a counterexample; these maximal
% certificates are an appropriate starting point for documenting
% both the counterexample and why it serves as a counterexample.

\begin{figure}
\lstinputlisting[linerange={pairing-end}]{code/fpcs.sig}
\lstinputlisting[linerange={pairing-end}]{code/fpcs.mod}
\caption{FPC for pairing}
\label{fig:pairing}
\end{figure}

>>>>

\begin{metanote}
  (AM): I would start with the logical reconstruction of \texttt{check}, adding the LJT-style proof terms~c\cite{Herbelin94,LengrandDM06} to fig \ref{fig:augmented} and then reverse engineer what we want/need to say about focusing and FPC. We \textbf{do} not talk about PBT yet, just logic programming (with proof terms). In sec 4 we package \texttt{check} in a PBT query and explain how \texttt{check} works for term generation.
\end{metanote}
\subsection{Focused proof systems}
\label{ssec:focused}

The proof search approach to encoding Horn clause computation
results in the structuring of proofs with repeated
switching between a \emph{goal-reduction} phase and a
\emph{backchaining} phase~\cite{miller91apal}.
%
The notion of \emph{focused proof systems}~\cite{andreoli92jlc,liang09tcs}
generalizes this view of proof construction by identifying the
following two phases. % of proof construction.
\begin{enumerate}
\item The \emph{negative}%
% \footnote{The terminology of negative and positive phases is a bit
%   unfortunate: historically, these terms do not refer to positive or negative
%   subformula occurrences but rather to certain semantic models used in
%   the study of linear logic~\cite{girard87tcs}.}
%
  phase corresponds to  goal-reduction: in this phase,
  inference rules that involve \emph{don't-care-non\-de\-ter\-min\-ism} are
  applied.
%
  As a result, there is no need to consider backtracking over choices
  made in building this phase.

@@

Further, if we view a particular FPC as a means of restricting proofs, it is
possible to build an FPC that \emph{restricts} proofs satisfying two FPCs
simultaneously.
%
% In particular, Figure~\ref{fig:pairing} defines an FPC based on the
% (infix) constructor \lsti{<c>}, which \emph{pairs} two terms of type
% \lsti{cert}.
%
The pairing experts for the certificate \lsti{Cert1 <c> Cert2} simply
request that the corresponding experts also succeed for both
\lsti{Cert1} and \lsti{Cert2} and, in the case of the \lsti{orE} and
\lsti{someE}, also return the same choice and substitution term,
respectively.
%
Thus, the query 
\begin{lstlisting}
?- check ((qheight 4) <c> (qsize 10)) B
\end{lstlisting}
will succeed if there is a proof of \lsti{B} that has a height less
than or equal to 4 while also being of size less than or equal to 10.
%
% A related use of the pairing of two proof certificates is to %use it to
% \emph{distill} or \emph{elaborate} proof certificates. 
% %
% For example, the proof certificate \lsti{(qsize 5 0)} is rather implicit
% since it will match any proof that used unfold exactly 5 times.
% %
% However, the query
% \begin{lstlisting}
% ?- check ((qsize 5 0) <c> (max Max)) B.
% \end{lstlisting}
% will store into the \lP variable \lsti{Max} more complete details of
% any proof that satisfies the \lsti{(qsize 5 0)} constraint.
% %
% In particular, this forms the infrastructure of 
%  an \emph{explanation} tool for attributing ``blame'' for
% the origin of a counterexample; these maximal
% certificates are an appropriate starting point for documenting
% both the counterexample and why it serves as a counterexample.

% \begin{figure}
% \lstinputlisting[linerange={pairing-end}]{code/fpcs.sig}
% \lstinputlisting[linerange={pairing-end}]{code/fpcs.mod}
% \caption{FPC for pairing}
% \label{fig:pairing}
% \end{figure}
